Author's second reply to the referee report of \\ \textsc{ ``PERSISTENCE IN FUNCTIONAL TOPOLOGY: EXISTENCE OF PERSISTENCE DIAGRAMS AND MORSE INEQUALITIES'' }

### 2022-10-12 sec/surfaces.tex
Remove double blank spaces
before referee 3

Having established topological conditions for the existence of persistence diagrams associated to filtrations and corresponding generalized Morse inequalities, we now describe how these results relate to Morse's general theory of functional topology \cite{Morse.1937, Morse.1938, Morse.1940}, and to its application in Morse and Tompkins' work on unstable minimal surfaces from \cite{Morse.1939}.
We will focus on the homological aspects of this approach, referring to \cite[Sections 4.3--5]{Bott.1980} for a general exposition.
For a more thorough presentation of the unstable minimal surface problem, including the analytical details see \cite[Section II.6]{Struwe.1988}, and for a historical account and an overview of subsequent results see \cite[Section 6]{Dierkes.2010} and \cite[Section 6.8.1]{Dierkes.2010b}.

@@

\begin{defi}[{{\cite[Definition II.6.1-II.6.2]{Struwe.1988}, \cite[p.~445, 466]{Morse.1939}}}]
\label{def:critical_set}
	Consider a real-valued function $F$ on a metric space $(M,d)$.
	A point $p \in M$ is called \emph{homotopically regular} if there exists a neighborhood $U$ of $p$ in $F_{\leq F(p)}$ and a continuous map $\varphi \colon U \times [0,1] \to M$, which satisfies $\varphi(\cdot, 0) = \operatorname{id}_U$ and $\varphi(p,1) \neq p$, such that for every compact subset $V \subseteq U$ there exists a \emph{displacement function} $\delta \colon \R_{\geq 0} \to \R_{\geq 0}$ with
	$\delta(e) = 0$ if and only if $e = 0$, and
	\[
	F(\varphi(x,s)) - F(\varphi(x,t)) \geq \delta(d(\varphi(x,s),\varphi(x,t)))
	\]
	for all $x \in V$ and $0 \leq s \leq t \leq 1$.
% 	\begin{itemize}
% 		\item $\delta(e) = 0$ if and only if $e = 0$, and
% 		\item $F(\varphi(x,s)) - F(\varphi(x,t)) > \delta(d(\varphi(x,s),\varphi(x,t)))$ for all $x \in V$ and $0 \leq s \leq t \leq 1$.
% 	\end{itemize}
	A point that is not homotopically regular is called \emph{homotopically critical}.
	Function values of homotopically critical points will be called \emph{critical values} and all other values will be called \emph{regular values}.
	A \emph{critical set} $S$ is a closed and open subset of the subspace of all homotopically critical points with a given function value.
	It is said to be of \emph{minimum type} if there exists a neighborhood $N$ of the closure $\overline{S}$ of $S$, taken in $M$, such that the function values on $N \setminus S$ strictly exceed the function value on $S$.
\end{defi}

@@

In \cite[p.~36]{Morse.1938}, the stronger notion of \emph{upper-reducibility}, without the prefix \emph{weakly}, is defined analogously to \cref{def:upper_reducible} with the slight difference that the existence of a displacement function is not only demanded for compact $V \subseteq U \cap F_{\geq c - \eta}$, but for all compact $V \subseteq U$.
Morse and Tompkins state the following Mountain Pass Theorem under slightly different assumptions, and with a slightly different conclusion.

@@

\begin{rem}\label{rem:critical_set}
	Our \cref{def:critical_set} of a critical set $S$ of \emph{minimum type} differs slightly from that of Morse and Tompkins \cite[p.~466]{Morse.1939}, who do not require the neighborhood $N$ of $S$ on which the function values exceed those on $S$ to contain the closure of $S$.
	Without this additional assumption, however, \cref{thm:mountain_pass} does not hold, as shown by the example $F \colon [0,1] \to \R$ with $F(0) = F(1) = 0$ and $F(t) = 1$ for $0 < t < 1$:
	$F$ has the four critical sets $\{0\}$, $\{1\}$, $\{0\} \cup \{1\}$ and $(0,1)$, which all satisfy the minimum type condition if the neighborhood $N$ need not contain their closure, but then there is no critical set that is not of minimum type.
\end{rem}

\begin{rem}\label{rem:mountain_pass_assumptions}
    In \cite[Corollary 7.1]{Morse.1939}, the assumptions that Morse and Tompkins use for $F$ and $M$ are that the sublevel set filtration is compact,  weakly upper-reducible, ``regular at infinity'', and that $M$ is ``locally $F$-connected''.
    Compactness is also required for our version, as is weak upper-reducibility. 
    Local $F$-connectedness is replaced by the $\HLC$ condition.
    (This point will be discussed in more detail in \cref{subsec:historic_hlc}).
    Regularity at infinity roughly corresponds to our assumption that $M$ is connected and the natural map $\colim \CH_0(F_{\leq \bullet}) \to \CH_0(M)$ is an isomorphism (see \cite[p.~444]{Morse.1939} and our proof of \cref{thm:unstable_minimial_surface}).
    This assumption may also replaced by the assumption that $\cp_0 = 1$, which we deduce from our assumptions in \cref{lem:essential_cech_dim}.
\end{rem}

\begin{rem}\label{rem:mountain_pass_conclusion}
    The precise statement of \cite[Corollary 7.1]{Morse.1939} postulates the existence of a homotopic critical point such that every critical set containing it has a positive first \emph{type number}.
    This corresponds to our statement in so far as the existence of a critical set not of minimum type implies that there are $\epsilon$ and $t$ such that $c_{1}^{\epsilon}(t) > 0$ (see \cref{lem:endpoint_implies_crit_pt} and the proof of \cref{thm:mountain_pass}).
    Note, however, that the existence of a positive cap number $c_{1}^{\epsilon}(t) > 0$ does not conversely imply the existence of a critical set not of minimum type:
    Consider the parabola $P = \{(x,x^2) \in \R^2 \mid x \in \R\}$, let $M = S^1 \times P$ and let $F \colon M \to \R$ be the Morse--Bott function given by the projection to the second coordinate in $P$.
    There is a single critical set $C = S^1 \times \{(0,0)\}$, which is of minimum type, but we still have $c_{1}^{\epsilon}(0) > 0$ for any $\epsilon > 0$ if $H$ is a homology theory such that $H_1(C) \neq 0$, e.g.\ singular or \v{C}ech homology.
    To establish the existence of a critical set not of minimum type in the proof of \cref{thm:mountain_pass}, we will show that not only $c_{1}^{\epsilon}(t) > 0$, but that we indeed have $\omega_1^{\epsilon}(t) > 0$, i.e., that the positive cap number is created by the death of a feature.
\end{rem}

\begin{rem}
	We mention that more general homology theories can be considered in \cref{thm:mountain_pass}.
	The precise hypotheses on the homology theory $\H$ are that it is additive, taking non-zero values on non-empty sets in dimension $0$, and such that $\H(F_{\leq \bullet})$ is continuous from above, i.e., $\H(F_{\leq s}) \to \lim_{s < t} \H(F_{\leq t})$ is an isomorphism for all~$s \in \R$.
	One may also replace the topological conditions of $M$ being connected, the natural map $\colim \H_0(F_{\leq \bullet}) \to \H_0(M)$ being an isomorphism, and $F$ being $\HLC$ by the algebraic conditions that $p_0 = 1$ and $\H(F_{\leq \bullet})$ be q-tame.
	% Recall that $\H$ is called \emph{additive} if $\H (A \sqcup B) \cong \H(A) \oplus \H(B)$ for all $A$ and $B$.
\end{rem}

@@

While more efficient and more general proofs for the existence of an unstable minimal surface (with respect to more natural topologies than $C^0$) have subsequently been established \cite{Struwe.1988,Dierkes.2010}, including less restrictive assumptions on the boundary curve, the original approach of Morse and Tompkins is notable
% for its close connection to persistent homology.
for its connections to other areas of mathematics.
As an illustration, we will now sketch a proof of \cref{thm:mountain_pass} using the previously developed machinery, starting with some intermediate results.
As a first step, we will give general formulas for \emph{numbers of births} $\alpha(t) = \sum_{q \in (t, \infty]} \m(t,q)$ and \emph{numbers of deaths} $\omega(t) = \sum_{p \in [-\infty, t)} \m(p,t)$ in q-tame persistence modules.

@@

We proceed by showing that function values with non-vanishing \emph{cap numbers} $c_d(t) \defeq \alpha_d(t) + \omega_{d-1}(t)$, for any $d$, are indeed critical values.
This corresponds to \cite[Theorem 8.1]{Morse.1938}.
Note that the definition of cap numbers used by Morse is phrased in terms of relative homology, differing from our definition in terms of absolute homology. The equivalence of both definition is shown in \cref{prop:cap_limits}.

\begin{lem}
\label{lem:endpoint_implies_crit_pt}
	Let $F \colon M \to \R$ be a weakly upper-reducible function on a metric space with compact sublevel set filtration.
	Assume that $\H(F_{\leq \bullet})$ is q-tame and continuous from above, and consider $t \in \R$ and its birth and death numbers $\alpha_d(t)$ and $\omega_{d-1}(t)$.
	If $\alpha_d(t) > 0$ or $\omega_{d-1}(t) > 0$, then $t$ is a critical value of $F$.
\end{lem}
\begin{proof}
	Following \cref{lem:sublevel_set_homotopy}, we know that if $t$ is a regular value, there exists $\epsilon > 0$ such that the inclusion $F_{\leq s} \hookrightarrow F_{\leq t}$ is a homotopy equivalence for all $s \in (t - \epsilon, t]$.
	Thus, $\H(F_{\leq \bullet})$ is continuous from below at every regular value.
	However, we assume $\H(F_{\leq \bullet})$ to be also continuous from above at every value, and in particular at regular values.
	Hence, \cref{cor:regular_value_no_endpoint} implies that $\alpha_d(t) = \omega_{d-1}(t) = 0$ whenever $t$ is regular, which proves the claim.
\end{proof}

@@

\begin{lem}
\label{lem:minimum_type_implies_some_birth_and_no_death}
	Let $F \colon M \to \R$ be a weakly upper-reducible function on a metric space with compact sublevel set filtration and let $S$ be a critical set of minimum type with value $t$.
	Assume that~$\H$~is additive and that $\H(F_{\leq \bullet})$ is q-tame and continuous from above.
	\begin{enumerate}
		\item The number of births at $t$ satisfies $\alpha_d(t) \geq \dim \H_d(S)$.
		\item If there are no homotopically critical points with value $t$ outside $S$, then the number of deaths at $t$ satisfies $\omega_d(t) = 0$ for all $d$.
	\end{enumerate}
\end{lem}
\begin{proof}
	We start by showing that $S$ is a topological summand of $F_{\leq t}$ in the sense that $F_{\leq t}$ is homeomorphic to the disjoint union $S \sqcup (F_{\leq t} \setminus S)$.
	It suffices to show that $S$ is open and closed in $F_{\leq t}$.
	By definition, there exists a neighborhood $N$ of $\overline{S}$ in $M$ such that the function values of $F$ on $N \setminus S$ exceed $t$.
	In particular, we have $F_{\leq t} \cap N = S$, showing that $S$ is open in $F_{\leq t}$.
	Because $N$ contains $\overline{S}$, we also obtain $F_{\leq t} \cap \overline{S} = S$, showing that $S$ is closed in $F_{\leq t}$.

@@

\begin{lem}
\label{lem:essential_cech_dim}
	Let $F \colon M \to \R$ be a function on a connected non-empty metric space with compact sublevel set filtration such that the natural map $\colim \CH_0(F_{\leq \bullet}) \to \CH_0(M)$ is an isomorphism and $\CH_0(F_{\leq \bullet})$ is q-tame.
	Then the essential dimension of $\CH_0(F_{\leq \bullet})$ satisfies $\cp_{0} = 1$.
\end{lem}
\begin{proof}
	Since $M$ is connected and non-empty, we have $\dim H_{0}(M) = 1$ for singular homology, and since $M$ is non-empty we have $\dim \CH_0(M) \geq 1$.
	Now the natural map from singular to \v{C}ech homology is always surjective for compact metric spaces in dimension 0 \cite{Eda.2000}, so $\dim \CH_0(M) = 1$.
	Together with our assumption on the colimit, we obtain $\cp_0 = \dim \colim \CH_0(F_{\leq \bullet}) = \dim \CH_0(M) = 1$.
\end{proof}

@@

	Now assume that $F$ has two distinct critical sets $S_{1}$ and $S_{2}$ of minimum type with values $t_{1}$ and $t_{2}$, respectively.
	Since both critical sets are non-empty, we have $\CH_{0}(S_{i}) \neq 0$, and thus the first assertion of \cref{lem:minimum_type_implies_some_birth_and_no_death} implies $\calpha_{0}(t_{i}) \geq \dim \CH_0(S_{i}) \geq 1$ for $i = 1,2$,
	indicating the existence of at least one feature for each $i = 1,2$ with birth $t_{i}$ and some positive persistence $\epsilon_i > 0$ in the persistence diagram of $\CH_0(F_{\leq \bullet})$.
	Choosing $0 < \epsilon < \min_i \epsilon_i$, this implies that
	$\calpha_{0}^{\epsilon}(t_{i}) \geq 1$ for $i = 1,2$,
	and thus
	$\cc_{0}^{\epsilon} \geq \calpha_{0}^{\epsilon}(t_{1}) + \calpha_{0}^{\epsilon}(t_{2}) \geq 1 + 1 = 2$.
	Now, from \cref{lem:essential_cech_dim} we obtain that $\cp_{0} = 1$, which yields $\comega_{0}^{\epsilon} = \cc_{0}^{\epsilon} - \cp_{0} \geq 2 - 1 = 1$.
	Thus, there must be some $t \in \R$ with $\comega_{0}(t) > 0$, so that we may apply \cref{lem:endpoint_implies_crit_pt} to obtain that the set $S$ of homotopically critical points at value $t$ is non-empty.
	If $S$ were of minimum type, then we would have $\comega_{0}(t) = 0$ by the second assertion of \cref{lem:minimum_type_implies_some_birth_and_no_death}, contradicting the choice of $t$.
	Hence, $S$ cannot be of minimum type, which finishes the proof.
\end{proof}

@@

Throughout his work on functional topology, Morse assumed slightly varying forms of local connectivity on the resulting sublevel set filtrations in order to obtain \mbox{q-tameness}.
In particular, Morse and Tompkins used the following condition from \cite{Morse.1938,Morse.1940} in their application to minimal surface theory:
\begin{displaycquote}[p.~431]{Morse.1940}
%\cite[p.~ 25]{Morse.1938}
	Let $p$ be a point of $M$ at which $F(p)=c$.
	The space $M$ is said to be \emph{locally $F$-connected} of order $r$ at~$p$ if corresponding to each positive constant $e$ there exists a positive constant $\delta$ such that each singular $r$-sphere on the $\delta$-neighborhood of $p$ and on $F_{c+\delta}$ bounds an $(r+1)$-cell of norm $e$ on $F_{c+e}$.
\end{displaycquote}
See also \cite[p.~ 25]{Morse.1938} and \cite[p.~464]{Morse.1939}, but note that the definitions given there contain evident typographical errors.
Expressed in similar language as \cref{s:connectivity}, the property of local $F$-connectedness of all orders is equivalent to the following notion, applicable to general topological spaces.

@@

\begin{proof}[Proof of \cref{thm:unstable_minimial_surface}]
    The Douglas functional $A_g$ is weakly upper-reducible by \cite[Theorem 5.1]{Morse.1939}, its sublevel set filtration is compact according to \cite[p.~448]{Morse.1939}, and as previously mentioned $\Omega_g$ is non-empty by \cite[p.~267-268]{Douglas.1931} because $g$ is rectifiable.
    Moreover, $\Omega_g$ is contractible and hence connected by \cite[Theorem 4.3]{Morse.1939}.
    The sublevel set filtration of $A_g$ is also $\HLC$ for \v{C}ech homology according to \cref{prop:douglas_hlc}.
    Finally, we have $\colim \CH_0((A_g)_{\bullet}) \to \CH_0(\Omega_g)$ is an isomorphism according to \cite[p.~444]{Morse.1939} because $A_g$ satisfies the \emph{regularity at infinity} condition by \cite[Theorem 4.3]{Morse.1939}.
    In total, \cref{thm:mountain_pass} applies to $A_g$.
    Since any homotopically critical point of $A_g$ corresponds to a minimal surface spanned by $g$ \cite[Theorem 6.2]{Morse.1939}, this implies the claim. 
\end{proof}

>>>>

Having established topological conditions for the existence of persistence diagrams associated to filtrations and corresponding generalized Morse inequalities, we now describe how these results relate to Morse's general theory of functional topology \cite{Morse.1937, Morse.1938, Morse.1940}, and to its application in Morse and Tompkins' work on unstable minimal surfaces from \cite{Morse.1939}.
We focus on the connection of homological and homotopical aspects of this approach, referring to \cite[Sections 4.3--5]{Bott.1980} for a general exposition.
For a more thorough presentation of the unstable minimal surface problem, including the analytical details, see \cite[Section II.6]{Struwe.1988}, and for a historical account and an overview of subsequent results see \cite[Section 6]{Dierkes.2010} and \cite[Section 6.8.1]{Dierkes.2010b}.

@@

\begin{defi}[{{\cite[Definition II.6.1-II.6.2]{Struwe.1988}, \cite[p.~445, 466]{Morse.1939}}}]
\label{def:critical_set}
	Consider a real-valued function $F$ on a metric space $(M,d)$.
	A point $p \in M$ is called \emph{homotopically regular} if there exists a neighborhood $U$ of $p$ in $F_{\leq F(p)}$ and a continuous map $\varphi \colon U \times [0,1] \to M$, which satisfies $\varphi(\cdot, 0) = \operatorname{id}_U$ and $\varphi(p,1) \neq p$, such that for every compact subset $V \subseteq U$ there exists a continuous \emph{displacement function} $\delta \colon \R_{\geq 0} \to \R_{\geq 0}$ with
	$\delta(e) = 0$ if and only if $e = 0$, and
	\[
	F(\varphi(x,s)) - F(\varphi(x,t)) \geq \delta(d(\varphi(x,s),\varphi(x,t)))
	\]
	for all $x \in V$ and $0 \leq s \leq t \leq 1$.
% 	\begin{itemize}
% 		\item $\delta(e) = 0$ if and only if $e = 0$, and
% 		\item $F(\varphi(x,s)) - F(\varphi(x,t)) > \delta(d(\varphi(x,s),\varphi(x,t)))$ for all $x \in V$ and $0 \leq s \leq t \leq 1$.
% 	\end{itemize}
	A point that is not homotopically regular is called \emph{homotopically critical}.
	Function values of homotopically critical points will be called \emph{critical values} and all other values will be called \emph{regular values}.
	A \emph{critical set} $S$ is a closed and open subset of the subspace of all homotopically critical points with a given function value.
	It is said to be of \emph{minimum type} if there exists a neighborhood $N$ of the closure $\overline{S}$ of $S$, taken in $M$, such that the function values on $N \setminus S$ strictly exceed the function value on $S$.
\end{defi}

@@

In \cite[p.~36]{Morse.1938}, the stronger notion of \emph{upper-reducibility}, without the prefix \emph{weakly}, is defined analogously to \cref{def:upper_reducible} with the slight difference that the existence of a displacement function is not only demanded for compact $V \subseteq U \cap F_{\geq c - \eta}$, but for all compact $V \subseteq U$.

Morse and Tompkins state the following Mountain Pass Theorem under slightly different assumptions, and with a slightly different conclusion.

@@

\begin{rem}\label{rem:critical_set}
    Our \cref{def:critical_set} of homotopically regular point differs slightly from that of \citet[p.~30]{Morse.1938} and \citet[p.~444--445]{Morse.1939} as their displacement functions are not explicitly assumed to be continuous.
    As used in the proof of \cref{lem:sublevel_set_homotopy}, we require this property to establish that a homotopically regular point with value $t$ admits a neighborhood $U$ and a homotopy $\varphi$ as in \cref{def:critical_set} such that for any compact $V \subseteq U$ there is $\epsilon > 0$ with $\varphi(V, 1) \subseteq F_{\leq t - \epsilon}$.
    That such neighborhoods exist is claimed by Morse in the proof of \cite[Lemma~8.1]{Morse.1938}, where continuity of the displacement functions is not mentioned explicitly.
    It is not clear to us whether the claim actually holds without continuity, or whether the assumption of continuity was made implicitly.
    Regardless, all displacement functions considered by \citet{Morse.1939} and used for the purposes of this work are indeed continuous.

	Our \cref{def:critical_set} of a critical set $S$ of \emph{minimum type} also differs slightly from that of Morse and Tompkins \cite[p.~466]{Morse.1939}, who do not require the neighborhood $N$ of $S$ on which the function values exceed those on $S$ to contain the closure of $S$.
	Without this additional assumption, however, \cref{thm:mountain_pass} does not hold, as shown by the example $F \colon [0,1] \to \R$ with $F(0) = F(1) = 0$ and $F(t) = 1$ for $0 < t < 1$:
	$F$ has the four critical sets $\{0\}$, $\{1\}$, $\{0\} \cup \{1\}$ and $(0,1)$, which all satisfy the minimum type condition if the neighborhood $N$ need not contain their closure, but then there is no critical set that is not of minimum type.
\end{rem}

\begin{rem}\label{rem:mountain_pass_assumptions}
	In \cite[Corollary 7.1]{Morse.1939}, the assumptions that Morse and Tompkins use for $F$ and $M$ are that the sublevel set filtration is compact,  weakly upper-reducible, ``regular at infinity'', and that $M$ is ``locally $F$-connected''.
	Compactness is also required for our version, as is weak upper-reducibility.
	Local $F$-connectedness is replaced by the $\HLC$ condition.
	(This point will be discussed in more detail in \cref{subsec:historic_hlc}).
	Regularity at infinity roughly corresponds to our assumption that $M$ is connected and the natural map $\colim \CH_0(F_{\leq \bullet}) \to \CH_0(M)$ is an isomorphism (see \cite[p.~444]{Morse.1939} and our proof of \cref{thm:unstable_minimial_surface}).
	This assumption may also be replaced by the assumption that $\cp_0 = 1$, which we deduce from our assumptions in \cref{lem:essential_cech_dim}.
\end{rem}

\begin{rem}\label{rem:mountain_pass_conclusion}
	The precise statement of \cite[Corollary 7.1]{Morse.1939} postulates the existence of a homotopic critical point such that every critical set containing it has a positive first \emph{type number}.
	This corresponds to our statement in so far as the existence of a critical set not of minimum type implies that there are $\epsilon$ and $t$ such that $c_{1}^{\epsilon}(t) > 0$ (see \cref{lem:endpoint_implies_crit_pt} and the proof of \cref{thm:mountain_pass}).
	Note, however, that the existence of a positive cap number $c_{1}^{\epsilon}(t) > 0$ does not conversely imply the existence of a critical set not of minimum type:
	Consider the parabola $P = \{(x,x^2) \in \R^2 \mid x \in \R\}$, let $M = S^1 \times P$ and let $F \colon M \to \R$ be the Morse--Bott function given by the projection to the second coordinate in $P$.
	There is a single critical set $C = S^1 \times \{(0,0)\}$, which is of minimum type, but we still have $c_{1}^{\epsilon}(0) > 0$ for any $\epsilon > 0$ if $H$ is a homology theory such that $H_1(C) \neq 0$, e.g.\ singular or \v{C}ech homology.
	To establish the existence of a critical set not of minimum type in the proof of \cref{thm:mountain_pass}, we will show that not only $c_{1}^{\epsilon}(t) > 0$, but that we indeed have $\omega_0^{\epsilon}(t) > 0$, i.e., that the positive cap number is created by the death of a feature.
\end{rem}

\begin{rem}
	We mention that more general homology theories can be considered in \cref{thm:mountain_pass}.
	The precise hypotheses on the homology theory $\H$ are that it is additive, taking non-zero values on non-empty sets in dimension $0$, and such that $\H(F_{\leq \bullet})$ is continuous from above, i.e., $\H(F_{\leq t}) \to \lim_{u > t} \H(F_{\leq u})$ is an isomorphism for all~$t \in \R$.
	One may also replace the topological conditions of $M$ being connected, the natural map $\colim \H_0(F_{\leq \bullet}) \to \H_0(M)$ being an isomorphism, and $F$ being $\HLC$ by the algebraic conditions that $p_0 = 1$ and $\H(F_{\leq \bullet})$ be q-tame.
	% Recall that $\H$ is called \emph{additive} if $\H (A \sqcup B) \cong \H(A) \oplus \H(B)$ for all $A$ and $B$.
\end{rem}

@@

While more efficient and more general proofs for the existence of an unstable minimal surface (with respect to more natural topologies on $\Omega_{g}$ than $C^0$) have subsequently been established \cite{Struwe.1988,Dierkes.2010}, including less restrictive assumptions on the boundary curve, the original approach of Morse and Tompkins is notable
% for its close connection to persistent homology.
for its connections to other areas of mathematics.
As an illustration, we will now sketch a proof of \cref{thm:mountain_pass} using the previously developed machinery, starting with some intermediate results.
As a first step, we will give general formulas for \emph{numbers of births} $\alpha(t) = \sum_{q \in (t, \infty]} \m(t,q)$ and \emph{numbers of deaths} $\omega(t) = \sum_{p \in [-\infty, t)} \m(p,t)$ in q-tame persistence modules.

@@

We proceed by showing that function values with non-vanishing \emph{cap numbers} $c_d(t) \defeq \alpha_d(t) + \omega_{d-1}(t)$, for any $d$, are indeed critical values.
This corresponds to \cite[Theorem 8.1]{Morse.1938}.
Note that the definitions of cap numbers used by Morse is phrased in terms of relative homology, differing from our definition in terms of absolute homology. The equivalence of both definition is shown in \cref{prop:cap_limits}.

\begin{lem}
\label{lem:endpoint_implies_crit_pt}
	Let $F \colon M \to \R$ be a weakly upper-reducible function on a metric space with compact sublevel set filtration.
	Assume that $\H(F_{\leq \bullet})$ is q-tame and continuous from above, and consider $t \in \R$ and its birth and death numbers $\alpha_d(t)$ and $\omega_{d-1}(t)$.
	If $\alpha_d(t) > 0$ or $\omega_{d-1}(t) > 0$ for some degree $d$, then $t$ is a critical value of $F$.
\end{lem}
\begin{proof}
	Following \cref{lem:sublevel_set_homotopy}, we know that if $t$ is a regular value, there exists $\epsilon > 0$ such that the inclusion $F_{\leq s} \hookrightarrow F_{\leq t}$ is a homotopy equivalence for all $s \in (t - \epsilon, t]$.
	Thus, $\H(F_{\leq \bullet})$ is continuous from below at every regular value.
	However, we assume $\H(F_{\leq \bullet})$ to be also continuous from above at every value, and in particular at regular values.
	Hence, \cref{cor:regular_value_no_endpoint} implies that $\alpha_d(t) = \omega_{d-1}(t) = 0$ for all $d$ whenever $t$ is regular, which proves the claim.
\end{proof}

@@

\begin{lem}
\label{lem:minimum_type_implies_some_birth_and_no_death}
	Let $F \colon M \to \R$ be a weakly upper-reducible function on a metric space with compact sublevel set filtration and let $S$ be a critical set of minimum type with value $t$.
	Assume that~$\H$~is additive and that $\H(F_{\leq \bullet})$ is q-tame and continuous from above.
	\begin{enumerate}
		\item The number of births at $t$ satisfies $\alpha_d(t) \geq \dim \H_d(S)$ for all $d$.
		\item If there are no homotopically critical points with value $t$ outside $S$, then the number of deaths at $t$ satisfies $\omega_d(t) = 0$ for all $d$.
	\end{enumerate}
\end{lem}
\begin{proof}
	We start by showing that $S$ is a topological summand of $F_{\leq t}$ in the sense that $F_{\leq t}$ is homeomorphic to the disjoint union $S \sqcup (F_{\leq t} \setminus S)$.
	It suffices to show that $S$ is open and closed in $F_{\leq t}$.
	By definition, there exists a neighborhood $N$ of $\overline{S}$ in $M$ such that the function values of $F$ on $N \setminus S$ exceed $t$.
	In particular, we have $F_{\leq t} \cap N = S$, showing that $S$ is open in $F_{\leq t}$.
	Because $N$ contains $\overline{S}$, we also obtain $F_{\leq t} \cap \overline{S} = S$, showing that $S$ is closed in $F_{\leq t}$.

@@

\begin{lem}
\label{lem:essential_cech_dim}
	Let $F \colon M \to \R$ be a function on a connected non-empty metric space with compact sublevel set filtration such that the natural map $\colim \CH_0(F_{\leq \bullet}) \to \CH_0(M)$ is an isomorphism and $\CH_0(F_{\leq \bullet})$ is q-tame.
	Then the essential dimension of $\CH_0(F_{\leq \bullet})$ satisfies $\cp_{0} = 1$.
\end{lem}
\begin{proof}
	Since $M$ is connected and non-empty, we have $\dim H_{0}(M) = 1$ for singular homology, and since $M$ is non-empty we have $\dim \CH_0(M) \geq 1$.
	Now the natural map from singular to \v{C}ech homology is always surjective for compact metric spaces in dimension 0 \cite{Eda.2000}, so the morphism $H_{0}(F_{\leq \bullet}) \to \CH_{0}(F_{\leq \bullet})$ is epi.
	Taking direct limits of vector spaces is exact, so the natural map $\colim H_{0}(F_{\leq \bullet}) \to \colim \CH_{0}(F_{\leq \bullet})$ is still epi. 
	We assume $\colim \CH_0(F_{\leq \bullet}) \to \CH_0(M)$ to be an isomorphism, so the composition $\colim H_0(F_{\leq \bullet}) \to \CH_0(M)$ is still epi.
	But this map factors through $H_{0}(M)$, which has dimension 1, so we get $\dim \CH_0(M) = 1$.
	Together with our assumption on the colimit, we obtain $\cp_0 = \dim \colim \CH_0(F_{\leq \bullet}) = \dim \CH_0(M) = 1$.
\end{proof}

@@

	Now assume that $F$ has two distinct critical sets $S_{1}$ and $S_{2}$ of minimum type with values $t_{1}$ and $t_{2}$, respectively.
	Since both critical sets are non-empty, we have $\CH_{0}(S_{i}) \neq 0$, and thus the first assertion of \cref{lem:minimum_type_implies_some_birth_and_no_death} implies $\calpha_{0}(t_{i}) \geq \dim \CH_0(S_{i}) \geq 1$ for $i = 1,2$,
	indicating the existence of at least one feature for each $i = 1,2$ with birth $t_{i}$ and some positive persistence $\epsilon_i > 0$ in the persistence diagram of $\CH_0(F_{\leq \bullet})$.
	Choosing $0 < \epsilon < \min_i \epsilon_i$, this implies that
	$\calpha_{0}^{\epsilon}(t_{i}) \geq 1$ for $i = 1,2$,
	and thus
	$\cc_{0}^{\epsilon} \geq \calpha_{0}^{\epsilon}(t_{1}) + \calpha_{0}^{\epsilon}(t_{2}) \geq 1 + 1 = 2$.
	Now, from \cref{lem:essential_cech_dim} we obtain that $\cp_{0} = 1$, which yields $\comega_{0}^{\epsilon} = \cc_{0}^{\epsilon} - \cp_{0} \geq 2 - 1 = 1$.
	Thus, there must be some $t \in \R$ with $\comega_{0}(t) \geq \comega^{\epsilon}_{0})(t) > 0$, so that we may apply \cref{lem:endpoint_implies_crit_pt} to obtain that the set $S$ of homotopically critical points at value $t$ is non-empty.
	If $S$ were of minimum type, then we would have $\comega_{0}(t) = 0$ by the second assertion of \cref{lem:minimum_type_implies_some_birth_and_no_death}, contradicting the choice of $t$.
	Hence, $S$ cannot be of minimum type, which finishes the proof.
\end{proof}

@@

Throughout his work on functional topology, Morse assumed slightly varying forms of local connectivity on the sublevel set filtrations of the functions he considered in order to obtain \mbox{q-tameness}.
In particular, Morse and Tompkins used the following condition from \cite{Morse.1938,Morse.1940} in their application to minimal surface theory:
\begin{displaycquote}[p.~431]{Morse.1940}
%\cite[p.~ 25]{Morse.1938}
	Let $p$ be a point of $M$ at which $F(p)=c$.
	The space $M$ is said to be \emph{locally $F$-connected} of order $r$ at~$p$ if corresponding to each positive constant $e$ there exists a positive constant $\delta$ such that each singular $r$-sphere on the $\delta$-neighborhood of $p$ and on $F_{c+\delta}$ bounds an $(r+1)$-cell of norm $e$ on $F_{c+e}$.
\end{displaycquote}
See also \cite[p.~ 25]{Morse.1938} and \cite[p.~464]{Morse.1939}, but note that the definitions given there contain evident typographical errors.
Expressed in similar language as \cref{defi:LHS}, the property of local $F$-connectedness of all orders is equivalent to the following notion, applicable to general topological spaces.

@@

\begin{proof}[Proof of \cref{thm:unstable_minimial_surface}]
    The Douglas functional $A_g$ is weakly upper-reducible by \cite[Theorem 5.1]{Morse.1939}, its sublevel set filtration is compact according to \cite[p.~448]{Morse.1939}, and as previously mentioned $\Omega_g$ is non-empty by \cite[p.~267-268]{Douglas.1931} because $g$ is rectifiable.
    Moreover, $\Omega_g$ is contractible and hence connected by \cite[Theorem 4.3]{Morse.1939}.
    The sublevel set filtration of $A_g$ is also $\HLC$ for \v{C}ech homology according to \cref{prop:douglas_hlc}.
    Finally, we have that $\colim \CH_0((A_g)_{\bullet}) \to \CH_0(\Omega_g)$ is an isomorphism according to \cite[p.~444]{Morse.1939} because $A_g$ satisfies the \emph{regularity at infinity} condition by \cite[Theorem 4.3]{Morse.1939}.
    In total, \cref{thm:mountain_pass} applies to $A_g$.
    Since any homotopically critical point of $A_g$ corresponds to a minimal surface spanned by $g$ \cite[Theorem 6.2]{Morse.1939}, this implies the claim.
\end{proof}
